Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2020-01-06 paper/manuscript.tex
Updates from Overleaf

\begin{abstract}
Alchemical free energy calculations are a useful tool for predicting free energy differences associated with the transfer of small molecules from one environment to another.
The hallmark of these methods is the use of bridging potential energy functions representing \emph{alchemical} intermediate states that cannot exist in chemistry. The data collected from these bridging alchemical thermodynamic states allows the computation of transfer free energies (or differences in transfer free energies) with orders of magnitude less simulation time than simulating the transfer process directly. 
While these methods are highly flexible, care must be taken in avoiding common pitfalls to ensure that computed free energy differences can be robust and reproducible for the chosen force field, and that appropriate corrections are included to permit direct comparison with experimental data.
In this paper, we review current best practices for several popular application domains of alchemical free energy calculations, including relative and absolute small molecule binding free energy calculations to biomolecular targets.
\todo[inline, color=green!20]{JDC: Should we migrate this document to \url{https://github.com/alchemistry}? ASJSM: Let's migrate with the submitted version, but leave it here for now?}
\end{abstract}

@@

\section{What are alchemical free energy methods?}
\label{sec:intro}
\todo[inline, color={green!20}]{@everyone: Any further input in this?}
Alchemical free energy (AFE) calculations compute free energy differences associated with transfer processes, such as the binding of a small molecule to a receptor or the transfer of a small molecule from an aqeuous to apolar phase~\cite{zwanzig1954hightemperature}. 
The calculations use non-physical intermediate states where atoms are changed from one type to another or can be removed or inserted by decoupling or coupling with their environment.

@@

In order to obtain a free energy from these AFE simulations, historically two common approaches were used for the analysis: free energy perturbation (FEP) using the Zwanzig relation~\cite{zwanzig1954hightemperature} or Bennett's acceptance ratio~\cite{bennett1976efficient} and thermodynamic integration (TI) for which the theory dates back decades but with the first computational applications emerging in the 1980's and 90's~\cite{kirkwood1935statistical, jorgensen1985monte, kollman1993free, wong1986dynamics, merz1989free}. Important studies continued during the 2000's often leading to improved implementation of the methods in common simulation software~\cite{vanderspoel2005gromacs, mermelstein2018fast, wang2015accurate, hedges2019biosimspace}. This foundational work, combined with the methodological, technological, and hardware improvements of the last 5-10 years have led to an explosion of interest.

As the field of molecular simulation moves towards millisecond timescales accurate AFE calculations on even more challenging problems will be in scope. In the meantime, today's users may find it hard to get started with these complex calculations whilst also keeping up with the fast pace of change. This best practice guide aims to serve as a go-to reference for current recommendations and tips for users of all experience.

@@

%%%%%%%%%%%%%%%%%%%%
% Prerequesites    %
%%%%%%%%%%%%%%%%%%%%
\section{Prerequisites and Scope}
\label{sec:pre}
%Here you would identify prerequisites/background knowledge that are assumed by your work and your checklist which you view as critical, ideally giving links to good sources on these topics.
%Checklists are normally focused on errors made by users with training and experience in molecular simulations, so you can assume a basic familiarity with the fundamentals of molecular simulations.
\todo[inline, color=green!20]{ASJSM: @everyone This section is up for review to also check flow with intro and rest of document.}
This guide is meant to serve as a comprehensive starting point for new practitioners, as well as a reference guide with a convenient checklist (Section~\ref{sec:checklist}) for more seasoned practitioners, to help with standard alchemical simulation and analysis practices. It can also help as a set of best practices which reviewers may wish to consider as they evaluate papers.

This guide aims to answer the following questions:
\begin{itemize}
    \item Is my problem suitable for an alchemical free energy calculation? 
    \item How do I choose the right free energy protocol and run it? \item How do I analyze my simulations accurately? 
    \item What software tools are available to simulate alchemical free energy protocols? 
\end{itemize}

@@

As some of the theoretical background can seem daunting, we provide an essentials guide to the theory behind alchemical free energy calculations in Section~\ref{sec:theory}.
We will cover topics that are key to the preparation~(\ref{sec:step1}), choice and use of correct protocols~(\ref{sec:step2}), and finally the best practices that should be used for the analysis of AFE calculations~(\ref{sec:step3}). 
Particular focus will be given to aspects of the molecular simulations which are unique to alchemical calculations---these include the calculation of transfer free energies (hydration free energies, partition coefficients, etc.), and binding free energies (absolute and relative).

While we try to address as many methods and practices as possible, the field of free energy calculations is broad, and there are many advanced topics that are left to future best practices documents focusing on specific issues. 
Below, we provide a non-exhaustive list of topics we have not addressed with some references to provide starting points on these more advanced topics:
\begin{itemize}
\item Covalent inhibition.
\item Complex association stoichiometry (beyond simple 1:1 association)~\cite{awesome reference}. 
\item Endpoint free energy methods such as MM-PBSA~\cite{genheden2015mm} and LIE~\cite{gutierrez-de-teran2012linear}
\item Free energy methods that extract the ligand using geometric order parameters and potential of mean force methods~\cite{heinzelmann2017attachpullrelease}
\item Forcefield dependence for protein, ligand, ions, co-solvents, and co-factors. Different studies have looked at the influence of force fields and it is assumed the user has made an adequate choice for the system under study.~\cite{loeffler2018reproducibility, vassetti2019assessment, lopes2015current} 
\end{itemize}

For convenience we have also compiled a list of common acronyms, common symbols used, and some essential terminology.
\begin{tcolorbox}[title=Acronyms, colback=blue!10!white]
    \textit{CPU} - Central Processing Unit\\
    \textit{AFE} - Alchemical Free Energy\\
    \textit{FEP} - Free Energy Perturbation\\
    \textit{GPCR} - G-Protein Coupled Receptor\\
    \textit{GPU} - Graphics Processing Unit\\
    \textit{MBAR} - Multistate Bennett Acceptance Ratio\\
    \textit{MCSS} - Maximum Common SubStructure\\
    \textit{MD} - Molecular Dynamics\\
    \textit{RBFE} - Relative Binding Free Energy\\
    \textit{RMSE} - Root Mean Square Error\\
    \textit{SAR} - Structure-Activity Relationships\\
    \textit{TI} - Thermodynamic Integration
\end{tcolorbox}
%\label{sec:abbreviations}
\begin{tcolorbox}[title=List of Symbols, colback=green!10!white]
U - Potential energy
\end{tcolorbox}

@@

\subsubsection*{The importance of the alchemical pathway\label{sec:important_path}}
\todo[inline, color={green!20}]{ASJM: This a bit waffly. Can we address: what do softcore potentials actually look like (equation)? Give an example of a lambda switching function(equation)? How should you chose your lambda intermediates? What is the best practice here? I have conveniently provided a figure to help with the discussion.}
\todo[inline, color={red!40}]{LNN: write common principles alchemical path choice}
\todo[inline, color={red!40}]{AR: write absolute-specific section on  alchemical path choice}
\todo[inline, color={red!40}]{BA: write relative-specific section on  alchemical path choice}

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig7_what_is_lam/Figure.pdf}
    \caption{By introducing a switching function that depends on a factor $\lambda$ it is possible to describe alchemical intermediates. In A it is shown that at $\lambda$ 0 the molecule is a fully interacting phenol and at $\lambda=1$ a fully interacting benzene. Intermediates can be modelled with a softcore potentials whose shape is illustrated in C with dependence on $\lambda$ and D with dependence on the potential energy. B shows an illustration of the probability distribution of the potential energies as the switching function takes values of $\lambda=0$ to $\lambda=1$, intermediates are required for a sufficient overlap in potential energies to estimate a free energy difference between $\lambda=0$ and $\lambda=1$. }
    \label{fig:fig_what_is_lambda}
\end{figure}

Both absolute and relative calculations must choose an alchemical pathway connecting initial and final states, which is in principle arbitrary but in practice affects the efficiency of the calculations considerably.
Some choices are particularly crucial -- for example, transformations involving insertions or deletions of atoms should employ soft-core potentials for Lennard-Jones or other hard-core interactions~\cite{}.
Other issues, such as whether absolute calculations retain intramolecular nonbonded interactions or remove these interactions, may be less critical and differ among studies in the literature~\cite{}.

Relative calculations introduce additional choices, including whether to define explicit intermediate states~\cite{} or leave these implicitly defined by the code~\cite{}.
Typically in single topology relative calculations it proves most efficient to first remove electrostatic interactions of any atoms which will be deleted, then modify other nonbonded interactions, then restore electrostatic interactions of any atoms which are being inserted.
Other schemes, such as simultaneously changing electrostatic and Lennard-Jones interactions, even with electrostatic ``soft core'' potentials, in our experience typically introduce errors and/or instabilities or are at least unreliable.
It is expected that similar considerations will apply to dual topology calculations, and the principle of first removing electrostatics and then removing steric interactions will likely serve well.

@@

A simple, but effective, way of checking for convergence is to compute the same property by multiple estimators. 
The uncertainty in the free energy, for example, has multiple ways to be estimated, e.g. through MBAR's estimator, through bootstrap, through multiple runs, etc. 
Independent of how the property is estimated, its important to remember that they are \textit{estimations of the property}, not the true underlying property itself. 
These estimators are usually consistent estimators, not necessarily unbiased ones though.
As such, it is a good idea to subject different estimators to the same data to see if they yield either the same estimate (within error and bias), or if they fluctuate wildly. 
This is not a perfect method as some estimators, such as Exponential Averaging, will converge significantly slower, or never, relative to more accurate estimators like MBAR. 
Therefore, it is a good idea to subject the estimators to different fractions of the data to see if the main estimator of free energy you have chosen is stable.

The determination of what information to collect during the simulation is paramount to its success. If, for instance, the free energy estimator selected is Thermodynamic Integration, but $\frac{dU}{d\lambda}$ information is not collected, then no amount of simulation time will provide a means to estimate the free energy. Once a combination of knowing what type of simulation you will run, which alchemical topology you will simulate, what alchemical path you will simulation along, and what your stopping conditions are, then you are ready to enumerate the information you should capture. Below is a sample of the minimal information you need for a set of common estimators:

@@

The potential derivative required for TI must generally be calculated during the simulation; it cannot be postprocessed by a code that does not evaluate the derivatives. 
If that option is unavailable, you can estimate it through finite difference (if sufficient information is collected), but understand this will introduce error and the BAR estimator may be a better, and simpler choice at that point as you will have at least the same level of information. 
The potential energy differences required for EXP, BAR, MBAR, and WHAM can be calculated either during the simulation or in post-processing. It is recommended to calculate the potential differences in code when possible to avoid extra overhead and possible errors produced by running the simulation twice, and to reduce the amount of stored information. 
Although TI must be usually be calculated in code, as it requires the derivative, there is one condition under which it actually has the fastest computation time. 
If the alchemical path you have chosen is a linear alchemical path, then you get $\frac{dU}{d\lambda} = U_0(\vec{x}) - U_1(\vec{x})$, which is the difference between the initial and final states. 
However, because of the problems with linear paths, this simplification is rarely particularly useful.

Free energy information should be saved as frequently as coordinate data, if not more frequently. 
The on-disk size of the data for free energy estimation is often significantly smaller than full atomic coordinates, so the information should be collected at least as frequently, if not more. 
However, the information should not be collected \textit{every} time step, as most free energy techniques are operated at equilibrium, and need equilibrated \textit{and decorrelated} samples for an unbiased estimate.
A sample collected every time step will likely result in most samples being discarded due to decorrelation routines in the analysis. However, if it is computationally cheap and disk space is plentiful, do save often. 
How decorelation impacts calculations, and how to compute it is discussed in other sections. [section ref]

@@

Once data has been collected from alchemical intermediates, it must be analyzed to produce an estimate of the free energy change (and its associated statistical uncertainty) for each leg of the thermodynamic cycle.
While a number of different estimators are available that will give consistent results under optimal circumstances, some approaches are recommended over others due to their robustness and ability to report on poor convergence.

@@

To minimize this effect, an initial portion of the simulation is often discarded to \emph{equilibration}~\cite{https://doi.org/10.33011/livecoms.1.1.5957}, with the idea of removing the most heavily biased initial portion of simulation data but retaining the unbiased \emph{production} region that represents a stationary Markov chain process sampling from the desired equilibrium target distribution.
Because the simulation time required for the atypical initial sampler state to relax toward equilibrium is a property of the specific system being simulated and the specific initial conditions selected, it is simplest to collect data for the whole process and use an automated algorithm to select how much data should be discarded to equilibration in a post-processing step.

@@

For additional discussion of working with correlated data and autocorrelation analysis, please refer to the work on Best Practices for Quantification of Uncertainty and Sampling Quality in Molecular Simulations~\cite{https://doi.org/10.33011/livecoms.1.1.5067}.

@@

The simplest method for calculating free energy
differences from simulations is the so-called \textit{Zwanzig
relationship}~\cite{zwanzig1954hightemperature}, also called one-sided exponential reweighting (EXP), or simply free energy perturbation, thought that term is sometimes used to encompass all ways of calculating free energy differences.

The free energy between two states defined by two different potential energy functions 
$U_0(\vec{q})$ and $U_1(\vec{q})$ over coordinate space $\vec{q}$ can be calculated as:
\begin{eqnarray}
\Delta f_{01} & = & \ln \expect{e^{-(U_1(\vec{q}) - U_0(\vec{q}))}}_0 =  \ln \expect{e^{-\Delta U(\vec{q})} }_0
\end{eqnarray}\label{eqn.zwanzig}
In words, we take the samples generated during our run with the potential energy function $U_0(\vec{q})$, recalculate what the energy difference would be to switch to potential energy function $U_1(\vec{q})$, and exponentially average the difference to get the free energy.  Here, the exponential average consists in multiplying by -1, exponentiating, averaging, taking the logarithm, and multiplying by -1 again.  This almost requires no extra code functionality to perform; one need only to save a full precision trajectory, and run an unmodified molecular simulation code using the $U_1$ in order to calculate the new energies of stored snapshots.  The analysis can be written in a line of code.  We note that this method is even more general, in that the instantaneous work to change the potential energy function from $U_0$ to $U_1$ can be replaced by the non-reversible work $W$ to make the same change some conditions~\cite{xx}.

>>>>

\begin{abstract}
Alchemical free energy calculations are a useful tool for predicting free energy differences associated with the transfer of small molecules from one environment to another.
The hallmark of these methods is the use of "bridging" potential energy functions representing \emph{alchemical} intermediate states that do not exist in as real chemical species. The data collected from these bridging alchemical thermodynamic states allows the computation of transfer free energies (or differences in transfer free energies) with orders of magnitude less simulation time than simulating the transfer process directly.

While these methods are highly flexible, care must be taken in avoiding common pitfalls to ensure that computed free energy differences can be robust and reproducible for the chosen force field, and that appropriate corrections are included to permit direct comparison with experimental data.

In this paper, we review current best practices for several popular application domains of alchemical free energy calculations, including relative and absolute small molecule binding free energy calculations to biomolecular targets.
\todo[inline, color=green!20]{JDC: Should we migrate this document to \url{https://github.com/alchemistry}? ASJSM: Let's migrate with the submitted version, but leave it here for now?}
\end{abstract}

@@

\section{What are alchemical free energy methods?}
\label{sec:intro}
\todo[inline, color={green!20}]{@everyone: Any further input in this?}
Alchemical free energy (AFE) calculations compute free energy differences associated with transfer processes, such as the binding of a small molecule to a receptor or the transfer of a small molecule from an aqeuous to apolar phase~\cite{zwanzig1954hightemperature}. 
The calculations use non-physical intermediate states where atoms are changed from one type to another or can be removed or inserted by turning on or off the interactions with their environment.

@@

In order to obtain a free energy from these AFE simulations, historically two common approaches were used for the analysis: free energy perturbation (FEP) using the Zwanzig relation~\cite{zwanzig1954hightemperature} or Bennett's acceptance ratio~\cite{bennett1976efficient} and thermodynamic integration (TI) for which the theory dates back decades but with the first computational applications emerging in the 1980's and 90's~\cite{kirkwood1935statistical, jorgensen1985monte, kollman1993free, wong1986dynamics, merz1989free}. Important studies continued during the 2000's often leading to improved implementation of the methods in common simulation software~\cite{vanderspoel2005gromacs, mermelstein2018fast, wang2015accurate, hedges2019biosimspace}. This foundational work, combined with the methodological, technological, and hardware improvements of the last 5-10 years have led to an explosion of interest and direct commercial application of the technologies~\cite{}.

As the field of molecular simulation moves towards millisecond timescales, accurate AFE calculations on even more challenging problems will become reasonable to perform. In the meantime, today's users may find it hard to get started with these complex calculations whilst also keeping up with the fast pace of change. This best practice guide aims to serve as a go-to reference for current recommendations and tips for users of all experience.

@@

%%%%%%%%%%%%%%%%%%%%
% Prerequesites    %
%%%%%%%%%%%%%%%%%%%%
\section{Prerequisites and Scope}
\label{sec:pre}
%Here you would identify prerequisites/background knowledge that are assumed by your work and your checklist which you view as critical, ideally giving links to good sources on these topics.
%Checklists are normally focused on errors made by users with training and experience in molecular simulations, so you can assume a basic familiarity with the fundamentals of molecular simulations.
\todo[inline, color=green!20]{ASJSM: @everyone This section is up for review to also check flow with intro and rest of document.}
This guide is meant to serve as a comprehensive starting point for new practitioners for free energy calculations.  It assumes moderate experience with molecular simulation packages and concepts, however.  It is also intended to serve as a reference guide with a convenient checklist (Section~\ref{sec:checklist}) for more seasoned practitioners, to help ensure their simulation meet standard alchemical simulation and analysis practices. It can also serve as a set of best practices to ensure simulation robustness and reproducibility which reviewers may wish to consider as they evaluate papers.

This guide aims to answer the following questions:
\begin{itemize}
    \item Is my problem suitable for an alchemical free energy calculation? 
    \item How do I choose the right free energy protocol and run it? 
    \item How do I analyze my simulations accurately? 
    \item What software tools are available to perform alchemical free energy calculations? 
\end{itemize}

@@

As some of the theoretical background can seem daunting, we provide an essentials guide to the theory behind alchemical free energy calculations in Section~\ref{sec:theory}.
In the remainder of this paper, we will cover topics that are key to the preparation~(\ref{sec:step1}), choice and use of correct protocols~(\ref{sec:step2}), and finally the best practices that should be used for the analysis of AFE calculations~(\ref{sec:step3}). 
Particular focus will be given to aspects of the molecular simulations which are unique to alchemical calculations---these include the calculation of transfer free energies (hydration free energies, partition coefficients, etc.), and binding free energies (absolute and relative).

While we try to address as many methods and practices as possible, the field of free energy calculations is broad, and there are many advanced topics that are left to future best practices documents focusing on specific issues. 
Below, we provide a non-exhaustive list of topics we have not addressed with some references to provide starting points on these more advanced topics:
\begin{itemize}
\item Covalent inhibition.
\item Free energies of mutation of protein side chains
\item nonspecific binding or multiple binding sites.
\item Complex association stoichiometry (beyond simple 1:1 association)~\cite{awesome reference}.
\item Endpoint free energy methods such as MM-PBSA~\cite{genheden2015mm} and LIE~\cite{gutierrez-de-teran2012linear}
\item Free energy methods that extract the ligand using geometric order parameters and potential of mean force methods~\cite{heinzelmann2017attachpullrelease}
\item Forcefield dependence for protein, ligand, ions, co-solvents, and co-factors. Different studies have looked at the influence of force fields and it is assumed the user has made an adequate choice for the system under study.~\cite{loeffler2018reproducibility, vassetti2019assessment, lopes2015current} 
\end{itemize}

For convenience we have also compiled a list of common acronyms, common symbols used, and some essential terminology.
\begin{tcolorbox}[title=Acronyms, colback=blue!10!white]
    \textit{CPU} - Central Processing Unit\\
    \textit{AFE} - Alchemical Free Energy\\
    \textit{BAR} - Bennett Acceptance Ratio\\
    \textit{FEP} - Free Energy Perturbation\\
    \textit{GPCR} - G-Protein Coupled Receptor\\
    \textit{GPU} - Graphics Processing Unit\\
    \textit{MBAR} - Multistate Bennett Acceptance Ratio\\
    \textit{MCSS} - Maximum Common Substructure\\
    \textit{MD} - Molecular Dynamics\\
    \textit{RBFE} - Relative Binding Free Energy\\
    \textit{RMSE} - Root Mean Square Error\\
    \textit{SAR} - Structure-Activity Relationships\\
    \textit{TI} - Thermodynamic Integration
\end{tcolorbox}
%\label{sec:abbreviations}
\begin{tcolorbox}[title=List of Symbols, colback=green!10!white]
U - Potential energy
\end{tcolorbox}

@@

\subsubsection*{The importance of the alchemical pathway\label{sec:important_path}}
\todo[inline, color={green!20}]{ASJM: This a bit waffly. Can we address: what do softcore potentials actually look like (equation)? Give an example of a lambda switching function(equation)? MRS: adressed these questions to some extent.  How should you chose your lambda intermediates? What is the best practice here? I have conveniently provided a figure to help with the discussion.}
\todo[inline, color={red!40}]{LNN: write common principles alchemical path choice. MRS: mostly wrote this, LNN can comment on it,}
\todo[inline, color={red!40}]{AR: write absolute-specific section on  alchemical path choice}
\todo[inline, color={red!40}]{BA: write relative-specific section on  alchemical path choice}

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig7_what_is_lam/Figure.pdf}
    \caption{Alchemical intermediates are createtd by making the potential energy depend on a an additional variable $\lambda$ that interpolates between the chemical endpoints. In A, at $\lambda=0$ the molecule is a fully interacting phenol and at $\lambda=1$,  a fully interacting benzene.  B shows an illustration of the probability distribution of the potential energies as the switching function takes values of $\lambda=0$ to $\lambda=1$, intermediates are required for a sufficient overlap in potential energies to estimate a free energy difference between $\lambda=0$ and $\lambda=1$.
    One of the most efficient families of intermediate pathways  are softcore potentials, whose shape is illustrated in C with dependence on $\lambda$ and D with dependence on the potential energy. Note how when $\lambda$ approaches 0, the energy smoothly approaches zero at all $r$, a necessary requirement for efficient calculations.  }
    \label{fig:fig_what_is_lambda}
\end{figure}

Both absolute and relative calculations must choose an alchemical pathway connecting initial and final states. In principle, because of the path independence of the free energy, any arbitrary pathway will give the correct free energy change, but the choice of pathway will greatly affect the efficiency of the calculations. 
Some choices are particularly crucial -- for example, transformations involving insertions or deletions of atoms should employ a path soft-core potential path for Lennard-Jones or other interactions with repulsive interactions that go to infinite energy at small radius~\cite{}.

The key consideration for choosing alchemical pathways is that the intermediate states that a given pathway produces should sample configurational ensembles that change as slowly as possible as $\lambda$ changes, while still managing to go from the initial state to the final state as $\lambda$ goes from 0 to 1

Another way of stating this is that intermediate states should sample molecular configurations that are as similar as possible to their neighboring states. The more similar the configurations are between intermediate states, the lower the statistical uncertainty is in the estimate of free energy between intervals.  This can be proven directly from the BAR and MBAR formulas~\citet{bennett1976efficient,Klimovich:JCAMD:2015}, though the exact same principles apply for TI.

The simplest possible alchemical pathway is a \textit{linear} pathway:

\begin{equation}
U(\vec{q},\lambda) = \lambda U_0(\vec{q}) + (1-\lambda)U_1(\vec{q})    
\end{equation}

so-called because the dependence on $\lambda$ is linear. This clearly satisfies the basic requirement that it gives the initial endpoint potential energy $U_0(\vec{q})$ when $\lambda=0$ and final endpoint energy $U_1(\vec{q})$ when $\lambda=1$.

For many energy terms, \textit{as long as a repulsive core remains on}, this is a very good approach.  For example, it can be shown that if van der Waals repulsions are left on, then the linear approach is very nearly the optimal path  possible for changing, removing, or inserting the electrostatic energy terms, with the path being within about \%10--20 of the minimum possible uncertainty~\cite{Naden:JCTC:2015} for a fixed amount of simulation, as well as being nearly optimally efficient for van der Waals attractive terms with repulsion terms turned on~\cite{Naden:JCTC:2014}. Although we are not aware of any quantitative tests for dipolar or higher multipole terms, theoretically it should behave equally well for those systems.

However, this approach ends up being terrible for removing or adding repulsive potentials that go to infinity quickly at or near the origin. One way to look at this is to examine how low $\lambda$ must go to reduce the energy at $0.5\sigma$ (the atomic size parameter) down to 1 $k_BT$, where thermal fluctuations make it possible for other atomic sites to penetrate routinely that deep. If we are trying go from a particle being present, and desire to make it disappear alchemically, then if the repulsive terms is of the form $\epsilon\frac{\sigma}{r}^{12}$, then if $\epsilon$ was 1 $K_BT$ at the temperature of interest, and we start with the particle present, then solving for $(1-\lambda)(1 kB_T){frac}{\sigma}{\frac{1}{2}\sigma}) = 1 k_B T$ we get $\lambda = 1-2^{-12} \sim 0.999976$.  So we have gone virtually all the way to the end of the transformation, there's still an impenetrable post in the middle of our simulation! This is not very much like the desired final state of no interactions between the particle and its environment. We can play around with a few ways of modifying this, like simulating many more intermediate states near $\lambda=1$. Various analyses have shown that this is not a very good strategy, however~\cite{Pham:JCP2011,Beutler:CPL:1994,Zacharias:JPC:1994,Blondel:JCC:2004}

What we need instead is some sort of function that smoothly gets rid of this infinity.  A large number of scheme have been tried~\cite{Beutler:CPL:1994,Zacharias:JPC:1994,Blondel:JCC:2004,Pham:JCP:2011,Pham:JCP:2011,Naden:JCTC:2014,Beutler:JCP,Donnini:JCC:2005,OTHERS!!!!}, but the most common strategy that appears to be the best practices is to use a "soft-core" potential, of the form:

\begin{equation}
    U(\vec{r_{ij}},\lambda) = 4\epsilon_{ij} \lambda \left(\frac{1}{(\alpha(1-\lambda) + (r_{ij}/\sigma_{ij})^6)^2} +  \frac{1}{\alpha(1-\lambda) + (r_{ij}/\sigma_{ij})^6}\right)\label{eq:softcore}
\end{equation}
Where $r_{ij}$ is the distance between two particles $i$ and $j$, $\epsilon_{ij}$ and $\sigma_{ij}$ are the Lennard-Jones parameters corresponding to the interaction between particles $i$ and $j$, and $\alpha$ is a constant (0.5 is optimal for the functional form shown above).  This functional form has exactly the property we are looking for: it recovers the Lennard-Jones potential when $\lambda=1$, and the other endpoint ($\lambda=0$), it is exactly zero for all $r_{ij}$ everywhere, and as $\lambda$ goes to zero, the  $\alpha(1-\lambda)$ term lowers the infinite energy in the core.  There are several different variants of the same functional form~\cite{Zacharias:JPC:1994,Beutler:CPL:1994,Pham:JCP:2011}, but the one in given in eq.~\ref{eq:softcore} is easy to understand and implement and fairly numerically stable. This functional form is shown in parts C and D of figure~\ref{fig:fig_what_is_lambda}.
It has been shown that more complicated forms are not significantly more efficient than eq.~\cite{Pham:JCP:2012}.  We recommend using the softcore potential given in eq.~\cite{eq:softcore}, unless there is a compelling reason otherwise.  Using a similar equation to eq.~\cite{eq:softcore} may be acceptable in most circumstances if that is what is supported in your chosen software. However, if you are inserting or removing entire atomic sites, we heavily recommend against using the linear approach; it will be very difficult to get correct results.

We have discussed optimal ways of disappearing or appearing Lennard-Jones interaction sites and turning on and off electrostatics terms.  What about performing both transformations at the same time? We can't turn off the electrostatics linearly at the same time we turn off the Lennard-Jones terms, as it would leave infinitely large attractive and repulsive electrostatic terms "bare" at small $\lambda$, resulting in the simulation crashing. It \textit{is} possible to apply the same soft core approach to the Coulomb interaction, but it turns out to be somewhat difficult to turn off two soft-core terms  terms simultaneously~\cite{Steinbrecher:JCC:2011}.  We therefore recommend performing the transformations in sequence; first, turning off all electrostatics for atoms that must be removed, inserting and removing Lennard-Jones sites (both the insertion and removal can be done simultaneously), and then turning electrostatics for the introduced particles on.  However, If there are no removals or introduction to atomic sites, then it is reasonable to change all interactions linearly.

Other issues, such as whether absolute calculations should retain or remove intramolecular nonbonded interactions, may be less critical and recommendations differ among studies in the literature~\cite{}, and reasonable efficiency can be often obtained with either choice even if some are somewhat better or worse than others. Our recommendation is to leave the intramolecular interactions on during the transformation for simplicity if there are no other known issues with this approach.  The key thing to watch out for is whether the total potential energy, and therefore the intermediate ensembles sampled, change smoothly from beginning to end.  If a configurational phase transition is noted in the process (like substantial dihedral rearrangement), then other steps must be taken, perhaps including modifying intramolecular interactions~\cite{???}.  These problems can be diagnosed by noticing lack of configuration space overlap between different simulations (see section "Are my simulations any good?)"

Relative calculations introduce additional choices, including whether to define explicit intermediate states~\cite{} or leave these implicitly defined by the code~\cite{}.
Typically in single topology relative calculations it proves most efficient to first remove electrostatic interactions of any atoms which will be deleted, then modify other nonbonded interactions, then restore electrostatic interactions of any atoms which are being inserted.
Other schemes, such as simultaneously changing electrostatic and Lennard-Jones interactions, even with electrostatic ``soft core'' potentials~\cite{Steinbrecher:JCC:2011}, in our experience typically introduce errors and/or instabilities or are at least unreliable.
It is expected that similar considerations will apply to dual topology calculations, and the principle of first removing electrostatics and then removing steric interactions will likely serve well.

@@

A simple, but effective, way of checking for convergence is to compute the same property by multiple estimators. 
The uncertainty in the free energy, for example, has multiple ways to be estimated, e.g. through standard error propagation methods (including MBAR's estimator, which is based on the same principles as standard error propagation), through bootstrap methods, through multiple independent runs, etc. 
Independent of how the property is estimated, its important to remember that they are \textit{estimations of the property}, not the true underlying property itself. 
These estimators are usually consistent estimators, meaning they will converge to the true answer in the limit of sufficient sampling, not necessarily unbiased ones though.
As such, it is a good idea to subject different estimators to the same data to see if they yield either the same estimate (within error and bias), or if they fluctuate wildly. 
This is not a perfect method as some estimators, such as exponential averaging, will converge significantly more slowely, relative to more accurate estimators like MBAR. 
Therefore, it is a good idea to apply the estimators to different fractions of the data to see if the main estimator of free energy you have chosen is stable.

Each method requries different data about the simulation be collected.  If, for instance, the free energy estimator selected is thermodynamic integration, then values of $\frac{dU}{d\lambda}$ at uncorrelated data points must be collected,  Once a combination of knowing what type of simulation you will run, which alchemical topology you will simulate, what alchemical path you will simulation along, and what your stopping conditions are, then you are ready to enumerate the information you should capture. Below is a sample of the minimal information you need for a set of common estimators:

@@

The potential derivative required for TI should generally be calculated during the simulation; only under every rare circumstances~\cite{Naden:JCTC:2015} can it postprocessed by a code that does not evaluate the derivatives. Many codes already have options for doing this.
If that option is unavailable, you can estimate it through finite difference (if sufficient information is collected), but this  will introduce significant error, and is generally not a best practice. Tthe BAR estimator may be a better, and simpler choice at that point as you will have at least the same level of information. 
The potential energy differences required for EXP, BAR, MBAR, and WHAM can be calculated either during the simulation or in post-processing. It is recommended to calculate the potential differences in code when possible to avoid extra overhead and possible errors produced by running the simulation twice, and to reduce the amount of stored information. 
Although TI must be usually be calculated in code, as it requires the derivative, there is one condition under which it actually has the fastest computation time. 
If the alchemical path you have chosen is a linear alchemical path, then you get $\frac{dU}{d\lambda} = U_0(\vec{x}) - U_1(\vec{x})$, which is the difference between the initial and final states. 
However, because of the problems with linear paths already discussed in this paper, this simplification is rarely that useful.

Free energy information should generally be saved more frequently than coordinate data, approximately at the rate that uncorrelated samples are produced.  
The on-disk size of the data for free energy estimation is often significantly smaller than full atomic coordinates, so the information should be collected frequently. 
However, the information should not be collected \textit{every} time step, as most free energy techniques are operated at equilibrium, and need equilibrated \textit{and decorrelated} samples for an unbiased estimate.
A sample collected every time step will likely result in most samples being discarded due to decorrelation routines in the analysis. However, if it is computationally cheap and disk space is plentiful, do save often. One may safely assume that the correlation time is greater than 100-200 fs even for relatively simple systems such as small molecules in solvent, so saving no more frequently than every 50-100 steps is recommended. 
How decorrelation impacts calculations, and how to compute it is discussed in other sections.  [section ref]

@@

Once data has been collected from alchemical intermediates, it must be analyzed to produce an estimate of the free energy change (and its associated statistical uncertainty) for each leg of the thermodynamic cycle.
While a number of different estimators are available that will give consistent results under optimal circumstances, some approaches are recommended over others due to their robustness and ability to provide information on poor convergence.

@@

To minimize this effect, an initial portion of the simulation is often discarded to \emph{equilibration}~\cite{Braun:L:2019}, with the idea of removing the most heavily biased initial portion of simulation data but retaining the unbiased \emph{production} region that represents a stationary Markov chain process sampling from the desired equilibrium target distribution.
Because the simulation time required for the atypical initial sampler state to relax toward equilibrium is a property of the specific system being simulated and the specific initial conditions selected, it is simplest to collect data for the whole process and use an automated algorithm to select how much data should be discarded to equilibration in a post-processing step.

@@

For additional discussion of working with correlated data and autocorrelation analysis, please refer to the work on Best Practices for Quantification of Uncertainty and Sampling Quality in Molecular Simulations~\cite{Grossfield:L:2019}.

@@

The simplest method for calculating free energy
differences from simulations is the so-called \textit{Zwanzig
relationship}~\cite{zwanzig1954hightemperature}, also called one-sided exponential reweighting (EXP), or simply free energy perturbation, thought this final term is sometimes used to encompass all ways of calculating free energy differences.

The (reduced) free energy difference $\Delta f_{01}$ between an initial state 0 and a final state 1 defined by two different potential energy functions 
$U_0(\vec{q})$ and $U_1(\vec{q})$ over coordinate space $\vec{q}$ can be calculated as:
\begin{eqnarray}
\Delta f_{01} & = & \ln \expect{e^{-(U_1(\vec{q}) - u_0(\vec{q}))}}_0 =  \ln \expect{e^{-\Delta u(\vec{q})} }_0
\end{eqnarray}\label{eqn.zwanzig}
and the average is over all samples from the simulation perfomed with $u_0$. In the case of NVT (canonical) sampling and assuming the masses do not change, then $u$ is just $U/k_BT$, and $f$ is $A/k_BT$, but it can be generalized to other ensembles with the proper definition of $f$ and $u$.
Described in words, we take the samples generated during our run with the potential energy function $U_0(\vec{q})$, recalculate what the energy difference would be if we switched to potential energy function $U_1(\vec{q})$, and average the exponential of the -1 times the energy difference to get -1 times the exponential of the free energy difference.   This almost requires no extra code functionality to perform; one need only to save a full precision trajectory, and run an unmodified molecular simulation code using the $U_1$ in order to calculate the new energies of stored snapshots.  The analysis can be written in a line of code.  We note that this method is even more general, in that the instantaneous work to change the potential energy function from $U_0$ to $U_1$ can be replaced by the non-reversible work $W$ to make the same change under the same equilibrium conditions at either end state~\cite{Jarzynski:PRL:1997,Jarzynski:APPB:1998,Crooks:PRE:2000}, though we do not go into all of the details of nonequilibrum transformations here, and refer the reader to more advanced treatments~\cite{Maragakis:JCP:2008,Oberhofer:JPCB:2005,Procacci:JCP:2015,Shirts:PRL:2003,Ytreberg:JCP:2004}
